---
title: "The IUCN RedList API"
format:
  html:
    toc: true
    number-sections: true
    code-fold: true
---

# Setup

We download the `iucnredlist` package from github using the package pacman. 

```{r}
#| label: setup
#| echo: false
#| quiet: true
#| results: false
#| message: false

pkgs <- c("DT",
          "here",
          "keyring",
          "tidyverse")

# install pacman if required
require("pacman")
pacman::p_load(pkgs, character.only = TRUE)

# obtain the redlist API package
pacman::p_install_gh("IUCN-UK/iucnredlist")
pacman::p_load("iucnredlist", character.only = TRUE)

```

Type `key_set("iucn-api-code")` into the console and paste your api key into the pop-up window. This will allow you to access data from the api without making the code available in your scripts etc.

# Obtain assessments

Using the API token we can obtain the RedLists of assessments by species.

## *Panthera leo*

The below example is for *Panthera leo* based on an `assessment_id`.

```{r}
#| label: test-api

# Initialize the API with your API key
api <- init_api(key_get("iucn-api-code"))

# Fetch assessment data for assessment_id = 266696959 (Panthera leo)
# Returns a list containing all assessment data
assessment_raw <- assessment_data(api, 266696959)

# Parse the 'raw' assessment data to tidy tibbles
assessment <- parse_assessment_data(assessment_raw)

# View e.g. threats or habitats
# The result is a tidy 'tibble()` of data
datatable(assessment$threats)
datatable(assessment$habitats)
```

## *Mastomys natalensis*

The API queries can also be run from a binomial name (i.e., *Mastomys natalensis*).

```{r}
# label: get-assessments-by-species

# Get all historic and latest assessments for Mastomys natalensis (not case-sensitive)
mnat <- assessments_by_name(api, genus = "mastomys", species = "natalensis")

datatable(mnat)
```

